Automated Verification for Real-Time Systems using Implicit Clocks and an Extended Antimirov Algorithm

### 2022-04-05 paper/tv.tex
m

%% Rights management information.  This information is sent to you
%% when you complete the rights form.  These commands have SAMPLE
%% values in them; it is your responsibility as an author to replace
%% the commands and values with those provided to you when you
%% complete the rights form.
\usepackage{enumerate}
\usepackage[utf8]{inputenc}
\usepackage{parcolumns}
\usepackage[spanish,english]{babel}
\usepackage{amsmath}
\usepackage{csquotes}
\usepackage{varwidth}
\newcommand{\key}[1]{\textcolor{purple}{\code{#1}}}
\newcommand{\jskey}[1]{\textcolor{blue}{\code{#1}}}
\newcommand{\es}{\textcolor{black}{\ensuremath{{es}}}}
\definecolor{palered-violet}{rgb}{0.86, 0.44, 0.58}
\definecolor{bittersweet}{rgb}{1.0, 0.44, 0.37}
\definecolor{brickred}{rgb}{0.8, 0.25, 0.33}
\definecolor{persianplum}{rgb}{0.44, 0.11, 0.11}
\newcommand{\hide}[1]{}
\newcommand{\siderule}[1]{
\code{\footnotesize{\textcolor{mGray}{#1}}}}
\usepackage{adjustbox}
\usepackage{ebproof}
\usepackage{cancel}
\newcommand{\env}{\code{\mathcal{E}}}

@@

\begin{figure}[]
  \vspace{-1mm}
  \renewcommand{\arraystretch}{1.1}
\small\begin{align*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%     Disjunction        %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
&\code{d, s, \varphi \models \effect_1 \vee \effect_2}  
&\m{iff}\quad & \code{d, s, \varphi \models \effect_1 } \ 
or\ \code{d, s, \varphi \models \effect_2 }  \\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%         Empty          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
&\code{d, s,  \varphi \models \pure{\pi}  , \empt}  
&\m{iff}\quad & \code{d{=}0}\ and  \ \code{\llbracket \pure{\pi} \rrbracket_s {=}  
\pure{True} } \ and\  \code{ \varphi {=} [] }\\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%         Event          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
&\code{d, s, \varphi \models \pure{\pi}  , ev }  
&\m{iff}\quad & \code{d{=}0}\ and  \ \code{\llbracket \pure{\pi} \rrbracket_s {=}  
\pure{True} }\ and \ \varphi {=} [ev]  \\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%   s  Concatenate        %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
&\code{d, s, \varphi \models \pure{\pi}, (\es_1\cdot\es_2)}  
&\m{iff}\quad & \exists \code{\varphi_1,\varphi_2}. \  
\code{ \varphi_1 {\lappend} \varphi_2 {=} \varphi}\ and\ 
\exists \code{d_1,d_2}. \  
\code{ d_1 {+} d_2 {=} d} \ 
  and \ \code{d_1, s, \varphi_1 {\models} \pure{\pi}  , \es_1 }
\ and \ \code{d_2, s, \varphi_2 {\models} \pure{\pi}  , \es_2 }\\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%    disjunction         %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
&\code{d, s, \varphi \models \pure{\pi}  , (\es_1\vee\es_2)}  
&\m{iff}\quad & \code{d, s, \varphi \models \pure{\pi}  , \es_1 } 
\ or\ \code{d, s, \varphi \models \pure{\pi}  , \es_2 } \\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%    t times         %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
&
\code{d, s, \varphi \models \pure{\pi}  ,\es\# t}  &\m{iff}\quad &
 \code{\llbracket \pure{\pi}  , t {=} 0 \rrbracket_{s} {=}  
 \pure{True} }, \ \code{d, s, \varphi \models \pure{\pi}  , \empt}\ or  \\
& & &  \code{\llbracket \pure{\pi}  , t {>} 0 \rrbracket_{s} {=}  
\pure{True} }, 
\  \exists \code{\varphi_1,\varphi_2}. \  \code{ \varphi_1 \lappend  \varphi_2 {=} \varphi}
\  and \ \code{d, s, \varphi_1 \models \pure{\pi}  , \es } \\
& & & \quad and \ \code{d, s, \varphi_2  \models (\pure{\pi}  , t {>} 0) \wedge  \es^{t\text{-}1} } \\
&\code{d, s, \varphi \models \pure{\pi}, \es^\star}  &\m{iff}\quad & there\ exist\ \code{\varphi_1,\varphi_2}
\   and \  \code{ \varphi_1 {+}{+} \varphi_2 {=} \varphi} 
 \  and \ \code{d, s, \varphi_1 \models \pure{\pi} , \es } 
 \   and \ \code{d, s, \varphi_2 \models \pure{\pi} , \es^\star }\\
& \code{d, s, \varphi \models false}  &\m{iff}\quad &
 \code{\llbracket \pure{\pi}  \rrbracket_{s} {=} \pure{False} } \ or \ \code{\varphi {=} \bott}
\end{align*}
  \vspace{-2mm}
\caption{Semantics of \timedEffects .}
\label{fig:Sementic}
\vspace{-2mm}
\end{figure}

@@

In this section, we give an axiomatic semantics model\footnote{Based on the operational semantics of Esterel and JavaScript's asynchrony formally defined in \cite{berry1992esterel} and  \cite{madsen2017model} respectively.} for the core language \code{\timedL}, by formalising a set of forward inductive rules. 
These rules transfer program states and systematically accumulate the effects syntactically. 
To define the model, we introduce an environment \env\ and describe a program state in a four-elements tuple \code{\langle \Pi, H, C, T \rangle}, with the following concrete domains:
\begin{flalign*}
\Pi {\defeq} t {\rightarrow} \pi, \qquad
H  {\defeq} es, \qquad
C {\defeq} \overrightarrow{\textbf{S} {\rightarrow} \Delta}\  | \ \textbf{S}?, \qquad
T {\defeq} t, \qquad 
\env {\defeq} \overrightarrow{\textbf{S}}, \qquad
\Delta {\defeq} \m{Present} \ | \ \m{Absent}  \ | \  \m{Undef}
\end{flalign*}

Let \code{\Pi} represents the pure constraints for all the terms; \code{H} represents the trace of \emph{history}; \code{C} represents the \emph{current} event or a waiting signal; 
\code{T} is a term binding \code{C}; 
\env\ be the environment containing all the local and output signals; 
\code{\Delta} marks the signal status.

@@

%\begin{lemma}
%The set of axioms (A1)-(A11) and (A15)-(A22) are complete for timed effects.
%\begin{proof} 
%
%\end{proof}
%\end{lemma}

 \begin{theorem}[Completeness]\label{Completeness}
 The set of axioms (A1)-(A22) is complete for \timedEffects.

%Given an inclusion \code{\effect_1 \CONTAIN \effect_2}, if \code{\effect_1 \CONTAIN \effect_2} is valid, then the TRS returns  \code{TRUE} by providing a proof.

\end{theorem}

\begin{proof} 
One can observe that by succinct applications of (A17) to (A22), we have a normal form, 
\begin{align}
\pi_1 : (I\# t_1) \cdot \es_1\ ||\ \pi_2 :  (J\# t_2) \cdot\es_2 \rightarrow  (\pi_1 \wedge \pi_2 \wedge t_1{=}t_2) :  ((I  {\cup} J) \# t_1) \cdot (\es_1\ ||\ \es_2)
\end{align}

Then for any ground \timedEffects \code{\es}, the expression \code{\epsilon || \es} can be reduced to either \code{\epsilon} or \code{\es} by the applications of (A15) and (A16).  
The expression \code{\es^\omega} can be reduced to either \code{\bot} or \code{\es^\omega} by the applications of (A12) to (A14).

Then, by the semantics defined in \figref{fig:Sementic}, effects constructed by Kleene star \code{\star} (possibly finite and possibly infinite) are supersets of effects constructed by \code{\omega} (definite infinite).

Then one can apply literally all the constructions of \cite{salomaa1966two} used in the completeness proof of \code{F_1} (A1) to (A11).

>>>>

%% Rights management information.  This information is sent to you
%% when you complete the rights form.  These commands have SAMPLE
%% values in them; it is your responsibility as an author to replace
%% the commands and values with those provided to you when you
%% complete the rights form.
\usepackage{enumerate}
\usepackage[utf8]{inputenc}
\usepackage{parcolumns}
\usepackage[spanish,english]{babel}
\usepackage{amsmath}
\usepackage{csquotes}
\usepackage{varwidth}
\newcommand{\key}[1]{\textcolor{purple}{\code{#1}}}
\newcommand{\jskey}[1]{\textcolor{blue}{\code{#1}}}
\newcommand{\es}{\textcolor{black}{\ensuremath{{es}}}}
\definecolor{palered-violet}{rgb}{0.86, 0.44, 0.58}
\definecolor{bittersweet}{rgb}{1.0, 0.44, 0.37}
\definecolor{brickred}{rgb}{0.8, 0.25, 0.33}
\definecolor{persianplum}{rgb}{0.44, 0.11, 0.11}
\newcommand{\hide}[1]{}
\newcommand{\siderule}[1]{
\code{\footnotesize{\textcolor{mGray}{#1}}}}
\usepackage{adjustbox}
\usepackage{ebproof}
\usepackage{cancel}
\newcommand{\env}{\code{\mathcal{S}}}

@@

\begin{figure}[]
  \vspace{-1mm}
  \renewcommand{\arraystretch}{1.2}
\small\begin{align*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%     Disjunction        %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
&\code{d, s, \varphi \models \effect_1 \vee \effect_2}  
&\m{iff}\quad & \code{d, s, \varphi \models \effect_1 } \ 
or\ \code{d, s, \varphi \models \effect_2 }  \\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%         Empty          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
&\code{d, s,  \varphi \models \pure{\pi}  , \empt}  
&\m{iff}\quad & \code{d{=}0}\ and  \ \code{\llbracket \pure{\pi} \rrbracket_s {=}  
\pure{True} } \ and\  \code{ \varphi {=} [] }\\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%         Event          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
&\code{d, s, \varphi \models \pure{\pi}  , ev }  
&\m{iff}\quad & \code{d{=}0}\ and  \ \code{\llbracket \pure{\pi} \rrbracket_s {=}  
\pure{True} }\ and \ \varphi {=} [ev]  \\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%     Concatenate        %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
&\code{d, s, \varphi \models \pure{\pi}, (\es_1\cdot\es_2)}  
&\m{iff}\quad & \exists \code{\varphi_1,\varphi_2}. \  
\code{ \varphi_1 {\lappend} \varphi_2 {=} \varphi}\ and\ 
\exists \code{d_1,d_2}. \  
\code{ d_1 {+} d_2 {=} d} \\
&&& \qquad  s.t. \ \code{d_1, s, \varphi_1 {\models} \pure{\pi}  , \es_1 }
\ and \ \code{d_2, s, \varphi_2 {\models} \pure{\pi}  , \es_2 }\\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%    disjunction         %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
&\code{d, s, \varphi \models \pure{\pi}  , (\es_1{\vee}\es_2)}  
&\m{iff}\quad & \code{d, s, \varphi \models \pure{\pi}  , \es_1 } 
\ or\ \code{d, s, \varphi \models \pure{\pi}  , \es_2 } \\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%    parallel         %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
&\code{d, s, \varphi \models \pure{\pi} , (ev_1{\cdot}\es_1)||(ev_2{\cdot}\es_2)}  
&\m{iff}\quad & \code{d, s, \varphi \models \pure{\pi}  , ev_1{\cdot} (\es_1||(ev_2{\cdot}\es_2)) } 
\ or\ \code{d, s, \varphi \models \pure{\pi}  , ev_2{\cdot}((ev_1{\cdot}\es_1)||es_2)  } \\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%    parallel  2        %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
&\code{d, s, \varphi \models \pure{\pi} , (ev\#t_1) || (\epsilon \# t_2)}  
&\m{iff}\quad & \code{d, s, \varphi \models \pure{(\pi{\wedge}t_1 {\geq}t_2)}, (ev\#t_1) } 
\ or\ \code{d, s, \varphi \models \pure{(\pi{\wedge}t_1 {<}t_2)}  , (ev\#t_1){\cdot}(\epsilon \# (t_2\text{-}t_1))  } \\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%    t times         %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
&\code{d, s, \varphi \models \pure{\pi}  ,\es\# t}  &\m{iff}\quad &
 \code{\llbracket \pure{\pi}  , t {=} 0 \rrbracket_{s} {=}  
 \pure{True} }, \ \code{d, s, \varphi \models \pure{\pi}  , \empt}\ or  \\
& & &  \code{\llbracket \pure{\pi}, t {>} 0 \rrbracket_{s} {=}  
\pure{True} }, 
\  \exists \code{\es_1,\es_2}. \  \code{ es_1 {\cdot}  es_2 {=} es}\\
&&& \qquad  s.t. \ \code{d, s, \varphi {\models} \pure{(\pi {\wedge}t_1{+} t_2{=}t)}  , 
(es_1\#t_1) {\cdot} (es_2\#t_2) } \\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%    star         %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
&\code{d, s, \varphi \models \pure{\pi}, \es^\star}  &\m{iff}\quad & 
\code{d, s, \varphi \models \pure{\pi}  , \empt}\ or \ 
\code{d, s, \varphi \models \pure{\pi}  , es {\cdot} \es^\star}\\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%    FALSE         %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
& \code{d, s, \varphi \models false}  &\m{iff}\quad &
 \code{\llbracket \pure{\pi}  \rrbracket_{s} {=} \pure{False} } \ or \ \code{\varphi {=} \bott}
\end{align*}
  \vspace{-2mm}
\caption{Semantics of \timedEffects .}
\label{fig:Sementic}
\vspace{-2mm}
\end{figure}

@@

In this section, we give an axiomatic semantics model\footnote{Based on the operational semantics of Esterel and JavaScript's asynchrony formally defined in \cite{berry1992esterel} and  \cite{madsen2017model} respectively.} for the core language \code{\timedL}, by formalising a set of forward inductive rules. 
These rules transfer program states and systematically accumulate the effects syntactically. 
To define the model, we introduce a stack \env\ and describe a program state in a four-elements tuple \code{\langle \Pi, H, T \rangle}, with the following concrete domains:
\begin{flalign*}
\Pi {\defeq} t {\rightarrow} \pi, \qquad\qquad
H  {\defeq} es, \qquad\qquad
T {\defeq} t, \qquad\qquad
\env {\defeq} \code{var {\rightarrow} val }
\end{flalign*}

Let \code{\Pi} represents the pure constraints for all the terms; \code{H} represents the trace of \emph{history}; \code{C} represents the \emph{current} event or a waiting signal; 
\code{T} is a term binding \code{C}; 
\env\ be the environment containing all the local and output signals;

@@
